Formal Modeling and Verification of a Rate-Monotonic Scheduling Implementation with Real-Time Maude

### 2016-03-24 RMS/submission.tex
modified related work partialy.

Considering schedulability test, Liu and
Layland~\cite{DBLP:journals/jacm/LiuL73} gave the famous sufficient
condition that a set of periodic tasks is schedulable with respect to
RMS if $\displaystyle \Sigma^n_{i=1} C_i/T_i \le n(2^{1/n}-1)$
holds. Then a more sufficient condition, known as \emph{Hyperbolic
  Bound}, which has the same complexity as Liu and Layland's, was
proposed in~\cite{DBLP:journals/tc/BiniBB03}. On the other hand,
necessary and sufficient conditions for schedulability were derived
independently in~\cite{DBLP:journals/rts/SpruntSL89}
and~\cite{audsley1993deadline}, requiring more sophisticated analysis
on the task set. Nevertheless, all these results take no overhead into
account, being not as realistic as ours. Katcher et al. did consider
overhead in their schedulability analysis, under several models based
on different kinds of popular
implementations~\cite{DBLP:journals/tse/KatcherAS93}.  However, our
target implementation is not in their scope.  Furthermore, compared
with those theoretical analyses, our approach based on formal modeling
and verification has a great advantage that if our schedulability test
answers ``no'', it returns at the same time a real counterexample,
which is able to guide our engineer to adjust the design.

@@

Condition~(iii). We must prove \deltaS{\deltaS{$s$}{$r$}}{$r'$} $=$
\deltaS{$s$}{$r+r'$} for all $r+r'\le$ \mteS{$s$}, with $s$ being a
system state of sort \verb|System|. Using the same symbols as in~(i),
we only consider the case $\mathit{ID}$\verb|::MaybeNat| in detail. 
By definition, the left side of the equation
\begin{IEEEeqnarray*}{Cl}
  & \deltaS{\deltaS{s}{r}}{r'}
\\  
= & \verb|(|\deltaTask{ID}{\deltaTask{ID}{L}{r}}{r'}
\\
  & \verb| |~T~STS~HW~\deltaIS{\deltaIS{ISRC}{r}}{r'}\verb|)|~\mbox{,}
\end{IEEEeqnarray*}
and the right side of the equation
\begin{IEEEeqnarray*}{Cl}
  & \deltaS{s}{r+r'}
\\  
= & \verb|(|\deltaTask{ID}{L}{r+r'}
\\
  & \verb| |~T~STS~HW~\deltaIS{ISRC}{r+r'}\verb|)|~\mbox{.}
\end{IEEEeqnarray*}
\deltaTask{$ID$}{\deltaTask{$ID$}{$L$}{$r$}}{$r'$} $=$
\deltaTask{$ID$}{$L$}{$r+r'$} holds by the associativity of $+$, while
\deltaIS{\deltaIS{$ISRC$}{$r$}}{$r'$} $=$ \deltaIS{$ISRC$}{$r+r'$}
holds since $(v$ \verb|monus| $r)$ \verb|monus| $r'$ $=$ $v$
\verb|monus| $(r+r')$ with $v$ of sort \verb|Time|. Thus
condition~(iii) holds.

@@

>>>>

Considering schedulability test, Liu and
Layland~\cite{DBLP:journals/jacm/LiuL73} gave the famous sufficient
condition that a set of periodic tasks is schedulable with respect to
RMS if $\displaystyle \Sigma^n_{i=1} C_i/T_i \le n(2^{1/n}-1)$
holds. Then a more sufficient condition, known as \emph{Hyperbolic
  Bound}, which has the same complexity as Liu and Layland's, was
proposed in~\cite{DBLP:journals/tc/BiniBB03}. On the other hand,
necessary and sufficient conditions for schedulability were derived
independently in~\cite{DBLP:journals/rts/SpruntSL89}
and~\cite{audsley1993deadline}, requiring more sophisticated analysis
on the task set. Nevertheless, all these results take no overhead into
account, being not as realistic as ours. Katcher et al. did consider
overhead in their schedulability analysis, under several models based
on different kinds of popular
implementations~\cite{DBLP:journals/tse/KatcherAS93}.  However, our
target implementation is not in their scope. Furthermore, compared
with those theoretical analyses, our approach based on formal modeling
and verification has three advantages. One is that if our
schedulability test answers ``no'', it returns at the same time a real
counterexample, which is able to guide our engineer to adjust the
design, by changing either the priorities or even the scheduling
algorithm. The second is that, when a fresh scheduling strategy is
applied, our analysis can be adjusted only by modifying the model,
while theoretical approaches may need thorough analyses and reasoning.
The last one is that, considering overhead and details of hardware do
introduce some kind of non-determinism into the model. For example, in
our model, if the running task is complete ``right'' at the time when
an interrupt request is performed, two different behaviors are
possible: (i) the system performs task switching, during which the
interrupt request is masked, hence the scheduling may be delayed; (ii)
the system answers the interrupt request immediately.

@@

Condition~(iii). We must prove \deltaS{\deltaS{$s$}{$r$}}{$r'$} $=$
\deltaS{$s$}{$r+r'$} for all $r+r'\le$ \mteS{$s$}, with $s$ being a
system state of sort \verb|System|. Using the same notations as in~(i),
we only consider the case $\mathit{ID}$\verb|::MaybeNat| in detail. 
By definition, the left side of the equation
\begin{IEEEeqnarray*}{Cl}
  & \deltaS{\deltaS{s}{r}}{r'}
\\  
= & \verb|(|\deltaTask{ID}{\deltaTask{ID}{L}{r}}{r'}
\\
  & \verb| |~T~STS~HW~\deltaIS{\deltaIS{ISRC}{r}}{r'}\verb|)|~\mbox{,}
\end{IEEEeqnarray*}
and the right side of the equation
\begin{IEEEeqnarray*}{Cl}
  & \deltaS{s}{r+r'}
\\  
= & \verb|(|\deltaTask{ID}{L}{r+r'}
\\
  & \verb| |~T~STS~HW~\deltaIS{ISRC}{r+r'}\verb|)|~\mbox{.}
\end{IEEEeqnarray*}
\deltaTask{$ID$}{\deltaTask{$ID$}{$L$}{$r$}}{$r'$} $=$
\deltaTask{$ID$}{$L$}{$r+r'$} holds by the associativity of $+$, while
\deltaIS{\deltaIS{$ISRC$}{$r$}}{$r'$} $=$ \deltaIS{$ISRC$}{$r+r'$}
holds since $(v$ \verb|monus| $r)$ \verb|monus| $r'$ $=$ $v$
\verb|monus| $(r+r')$ with $v$ of sort \verb|Time|. Thus
condition~(iii) holds.

@@

  About non-determinism: task finishes right at the time interrupt
  comes, theoretical results cannot solve
